---
title: "Analyzing 30 Rock Data"
author: "Phillip Sanderell"
output: github_document
---

"Science is my most favorite subject, especially the Old Testament." Continuing my one-show streak of analyzing data for my favorite television shows, I have chosen to explore Tina Fey's *30 Rock*. In this post I look forward to learning about what it takes to make a truly fantastic show, and perhaps impress my friends from The Second City.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(showtext)
showtext_auto()
font_add("Monsterrat", "Montserrat-Regular.ttf")
yellow <- "#FFFFAC"
light_theme <- 
  theme_minimal() +
  theme(text = element_text(family = "Monsterrat", color = "black"),
        plot.title = element_text(size = 15),
        axis.text = element_text(color = "black", size = 10),
        axis.title = element_text(color = "black", size = 12),
        plot.title.position = "plot")
theme_set(light_theme)
```

```{r include=FALSE}
rock <- read_csv("https://raw.githubusercontent.com/philliphresh/30-rock-2021/main/rock.csv")
```

# Exploring episode metadata

## IMDb rating

IMDb ratings tend to be my go-to metric for scoring episodes, and something that I find particularly interesting. I suppose I assume that episodes with higher IMDb ratings had higher Neilsen ratings or generated more audience for the show. The plot below shows the distribution of IMDb ratings for every regular-season episode of the show.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, dpi=900}
nbc_pallete <- c(`1` = "#FCB711",
                 `2` = "#F37021",
                 `3` = "#CC004C",
                 `4` = "#6460AA",
                 `5` = "#0089D0",
                 `6` = "#0DB14B",
                 `7` = "black")

rock %>% 
  mutate(season = as.character(season)) %>% 
  ggplot(aes(rating, fill = season)) +
  geom_histogram(binwidth = .1) +
  scale_fill_manual(values = nbc_pallete,
                    name = "Season") +
  theme(panel.grid.minor.y = element_blank(),
        plot.caption.position = "plot") +
  labs(title = "30 ROCK's IMDb ratings generally range between 7.5 and 8.6",
       x = "Episode Rating",
       y = "Count",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

The ratings do seem to have an approximately normal distribution centered at eight stars. Something worth noting is that while IMDb ratings range from 0 to 10, I find episode ratings for various shows hover around 8 starts---perhaps there is some psychological reason for that, but unfortunately we have no way of knowing, medicine is not a science.

The plot above is useful for examining episode ratings overall, but it is easy to tell if certain seasons rate better than others. The next plot was made in an effort to make this more clear. Each dot represents an episode going in order of air date from left to right, colored by season, and sized by number of IMDb votes. The "notched" box plots show the 25th, 50th, and 75th percentiles of ratings for each season---if the notches in different season's boxes don't overlap, it suggests the medians are significantly different.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, dpi=900}
season_labels <- c("Season 1", "2", "3", "4", "5", "6", "7")
names(season_labels) <- 1:7

ann_1 <- tibble(season = 1, ep_no = 7, rating = 7.3, hjust = 0,
                label = "Pilot")
ann_2 <- tibble(season = 1, ep_no = 3, rating = 7.4, hjust = 0, 
                label = "Unaired Pilot")
ann_3 <- tibble(season = 6, ep_no = 16, rating = 7.3, hjust = 1, 
                label = "Queen of Jordan 2: The Mystery\nof the Phantom Pooper")
ann_4 <- tibble(season = 7, ep_no = 10, rating = 9.0, hjust = 1, 
                label = "Hogcock!/Last Lunch")

rock %>% 
  # arrange(-rating)
  mutate(season = fct_reorder(as.character(season), air_date)) %>% 
  
  ggplot(aes(ep_no, rating)) +
  geom_point(aes(size = no_votes, color = season)) + 
  geom_boxplot(aes(y = rating), color = "black",
               size = 1, alpha = .2, notch = TRUE, 
               coef = 0, outlier.color = NA) +
  facet_grid(cols = vars(season), 
             scales = "free_x", 
             labeller = labeller(season = season_labels)) +
  geom_text(aes(label = label, hjust = hjust), data = ann_1) +
  geom_text(aes(label = label, hjust = hjust), data = ann_2) +
  geom_text(aes(label = label, hjust = hjust), data = ann_3, lineheight = .7) +
  geom_text(aes(label = label, hjust = hjust), data = ann_4) +
  scale_color_manual(values = nbc_pallete) +
  scale_fill_manual(values = nbc_pallete) +
  guides(size = "none", color = "none", fill = "none") +
  coord_cartesian(clip = 'off') +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        strip.text = element_text(size = 10),
        strip.background = element_rect(fill = "grey95", size = 0)) +
  labs(title = "The distribution of ratings gets more variable over time",
       y = "Episode Rating",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

Based on the notched box plots, season six had a significantly lower median rating than seasons two and three---perhaps that explains why season seven is the last with only 12 episodes. NBC is quick to cancel non-cop shows when they're not doing well.

## Prevalence of writers

The plot below lists every writer of the show in order of total episodes written for the series. The median IMDb rating for each writer's episodes is listed beside their name. Please note, on the full cast and crew page for each episode, IMDb lists some people in the writing credits section as 'story editor', 'executive story editor', etc., and it always lists Tina Fey as 'created by'. The counts in the plot below only capture a writer if they are listed as 'written by'.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=7, dpi=900}
rock %>% 
  separate(writers, paste0("writer_", 1:4), sep = "\\_\\_") %>% 
  pivot_longer(writer_1:writer_4, names_to = NULL, values_to = "writer") %>% 
  filter(!is.na(writer)) %>% 
  group_by(writer) %>% 
  mutate(tot_eps = n(),
         med_rating = median(rating)) %>% 
  group_by(season, writer) %>% 
  summarise(n = n(),
            tot_eps = mean(tot_eps),
            med_rating = mean(med_rating)) %>% 
  ungroup() %>% 
  mutate(writer2 = paste0(writer, "   (",
                          scales::comma(med_rating, .1),
                          ")")) %>% 
  mutate(writer2 = fct_reorder(writer2, tot_eps)) %>% 
  
  ggplot(aes(n, writer2, fill = as.character(season))) + 
  geom_col() + 
  facet_grid(cols = vars(season),
             labeller = labeller(season = season_labels)) +
  scale_fill_manual(values = nbc_pallete) +
  scale_x_continuous(breaks = seq(1, 9, 2)) +
  guides(fill = "none") +
  theme(panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_rect(fill = "grey95", size = 0)) +
  labs(title = "There seem to be three tiers of writers:",
       subtitle = "    A.    Original writing team\n    B.    New writers brought on in season four\n    C.    Various guest writers",
       x = "Number of episodes written",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

Tina Fey clearly did not sleep, having had the highest number of episodes written (or tied for) in five out of seven seasons, all while staring in every episode. Robert Carlock wrote the second most total episodes for the show, which is fortunate because I love his writing style: quick, ridiculous, and too witty to keep up with sometimes. Tina Fey and Robert Carlock went on to co-create *The Unbreakable Kimmy Schmidt*, which would be another interesting show to analyze, bringing along a lot of 30 Rock writing alumna. Tracy Wigfield created the NBC series *Great News*, another great watch. I believe Donald Glover is credited as a story consultant writer on some episodes that do not show up here, but looks like he worked on the actual script for two episodes.

### Did bringing on more writers cause the ratings downfall?

It's a leading question, I know, but it is curious to see that 30 Rock's ratings started going downhill in season four and that happens to also be the season that the main writing staff added about four new writers.

I think a way to answer this question simply is to compare the distribution of ratings in the fourth through seventh seasons between the original writing team and the new writers (excluding guest writers). If they have similar distributions we cannot make any conclusions about the new writers, if the new writers have lower ratings then we can reasonably assume they did not do the show any favors.

For the purposes of this question, the original writing team is:

-   Tina Fey

-   Robert Carlock

-   Jack Burditt

-   Matt Hubbard

-   Kay Cannon

-   Ron Weiner

-   John Riggi

The new writers are:

-   Tracey Wigfield

-   Josh Siegal

-   Dylan Morgan

-   Vali Chandrasekaran

-   Tom Ceraulo

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, dpi=900}
rock %>% 
  separate(writers, paste0("writer_", 1:4), sep = "\\_\\_") %>% 
  pivot_longer(writer_1:writer_4, names_to = NULL, values_to = "writer") %>% 
  filter(!is.na(writer)) %>% 
  mutate(writing_team = case_when(writer %in% c("Tina Fey",
                                                "Robert Carlock",
                                                "Jack Burditt",
                                                "Matt Hubbard",
                                                "Kay Cannon",
                                                "Ron Weiner",
                                                "John Riggi") ~ "Original Writing Team",
                                  writer %in% c("Josh Siegal",
                                                "Dylan Morgan",
                                                "Vali Chandrasekaran",
                                                "Tracey Wigfield",
                                                "Tom Ceraulo") ~ "New Writers",
                                  TRUE ~ "Guest Writers")) %>% 
  mutate(writing_team = fct_relevel(writing_team, "Original Writing Team", "New Writers", "Guest Writers")) %>% 
  mutate(color = case_when(writing_team == "Original Writing Team" ~ "black",
                           writing_team == "New Writers" ~ "#FCB711",
                           writing_team == "Guest Writers" ~ "grey50")) %>% 
  mutate(season_labels = if_else(season == 1, "Season 1", season %>% as.character)) %>% 
  
  ggplot(aes(writing_team, rating)) +
  geom_boxplot(aes(color = color), coef = 0, outlier.alpha = 0) +
  facet_grid(cols = vars(season), 
             labeller = labeller(season = season_labels)) +
  scale_color_identity(guide = "legend",
                       name = NULL,
                       limits = c("black", "#FCB711", "grey50"),
                       labels = c("Original Writing Team", "New Writers", "Guest Writers")) +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.direction = "horizontal",
        legend.position = "bottom",
        strip.background = element_rect(fill = "grey95", size = 0)) +
  labs(title = "New writers have slightly lower ratings than the original writing team\nfrom season four on",
       subtitle = "Guest writers are less predictible",
       y = "Episode Rating",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

## Prevalence of directors

The plot below lists every director of the show in order of total episodes directed for the series. The median IMDb rating for each director's episodes is listed beside their name.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=7, dpi=900}
rock %>% 
  separate(directors, paste0("director_", 1:4), sep = "\\_\\_") %>% 
  pivot_longer(director_1:director_4, names_to = NULL, values_to = "director") %>% 
  filter(!is.na(director)) %>% 
  group_by(director) %>% 
  mutate(tot_eps = n(),
         med_rating = median(rating)) %>% 
  group_by(season, director) %>% 
  summarise(n = n(),
            tot_eps = mean(tot_eps),
            med_rating = mean(med_rating)) %>% 
  ungroup() %>% 
  mutate(director2 = paste0(director, "   (",
                            scales::comma(med_rating, .1),
                            ")")) %>% 
  mutate(director2 = fct_reorder(director2, tot_eps)) %>% 
  
  ggplot(aes(n, director2, fill = as.character(season))) + 
  geom_col() + 
  facet_grid(cols = vars(season),
             labeller = labeller(season = season_labels)) +
  scale_fill_manual(values = nbc_pallete) +
  scale_x_continuous(breaks = seq(1, 9, 2)) +
  guides(fill = "none") +
  theme(panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_rect(fill = "grey95", size = 0)) +
  labs(title = "The top four directors dominate, others seem to come when available",
       x = "Number of episodes directed",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

I wonder what happened to Adam Bernstein? He directed just as much as Don Scardino in season one and then didn't direct anymore. Jeff Richmond directed a number of later episodes, avid watchers may recognize his name as the composer for the series (or as Tina Fey's husband in real life). I notice Robert Carlock, the second-most credited writer, directed an episode in the final season---I wonder if there is any other writer/director overlap.

## Examining characters listed in episode description

Can we find any relationships with character names and IMDb ratings?

### A little more web scraping needs performed to gather a list of character names

```{r, include=FALSE}
library(rvest)

full_credits_link <- "https://www.imdb.com/title/tt0496424/fullcredits/?ref_=tt_cl_sm"

full_credits_page <- read_html(full_credits_link)

# Get list of character names
full_credits <- 
  full_credits_page %>% 
  html_elements(".character a:nth-child(1) , .character a:nth-child(1) font") %>% 
  html_text(trim = TRUE) %>% 
  # Limit ot top 50 characters
  .[1:50] %>% 
  tibble(full_name = .) %>% 
  mutate(first_name = word(full_name, 1)) %>% 
  # Fix some of the tricky names
  mutate(first_name = case_when(full_name == "Dot Com" ~ "Dot Com",
                                full_name == "J.D. Lutz" ~ "Lutz",
                                full_name == "Dr. Leo Spaceman" ~ "Dr. Spaceman",
                                full_name == "Stage Manager" ~ "Stage Manager",
                                full_name == "Young Liz" ~ "Young Liz",
                                full_name == "Dr. Drew Baird" ~ "Dr. Baird",
                                full_name == "Old Janitor" ~ "Old Janitor",
                                TRUE ~ first_name))

# Just making sure there are no repeats
full_credits %>% 
  count(first_name)
```

View the README.Rmd file in this Github repo to see the extra web scraping code to gather character names.

### Plotting character rating prevalence

I'm interested in seeing if certain characters being the main focus of an episode drives the IMDb rating up or down. Unfortunately a large assumption needs to be made: characters being the main focus of an episode are listed in the episode description on IMDb. Of course I'm sure this assumption is not perfect, there are certainly episode descriptions that don't mention characters with prominent story lines in the episode.

The plot below shows the top 30 or so characters listed in episode descriptions, with their 25th, 50th, and 75th percentiles of episode ratings given on the left bar, dot, and right bar of each row. The larger the dot, the more episodes with that character mentioned in the description---the top 5 characters are highlighted.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6, dpi=900}
# Cross reference character names with episode descriptions
characters <- rock
for (i in full_credits$first_name) {
  characters <- 
    characters %>% 
    mutate(str_detect(ep_desc, i)) %>% 
    rename(!!enquo(i) := `str_detect(ep_desc, i)`)
}

characters %>% 
  pivot_longer(Liz:Lester, 
               names_to = "character", 
               values_to = "in_episode_desc") %>% 
  filter(in_episode_desc == TRUE) %>% 
  group_by(character) %>% 
  mutate(n = n()) %>% 
  slice_max(n, n = 20) %>% 
  summarise(n = mean(n),
            p25 = quantile(rating, .25),
            p50 = quantile(rating, .50),
            p75 = quantile(rating, .75)) %>% 
  mutate(character = fct_reorder(character, n)) %>% 
  mutate(character = fct_reorder(character, p50)) %>%
  
  ggplot(aes(y = character)) +
  geom_point(aes(x = p50, size = n)) +
  geom_errorbar(aes(xmin = p25, xmax = p75)) +
  gghighlight::gghighlight(character %in% c("Jack", "Liz", "Tracy", "Jenna", "Kenneth"), 
                           use_direct_label = FALSE,
                           unhighlighted_params = aes(color = "grey50")) +
  
  annotate("text", 7.47, 3.5, hjust = 1, 
           label = "25th") +
  geom_curve(aes(x = 7.46, y = 3, xend = 7.49, yend = "Hazel"), arrow = arrow(length = unit(0.01, "npc"))) +
  annotate("text", 7.6, 3.5, hjust = .5, 
           label = "50th") +
  geom_curve(aes(x = 7.6, y = 3, xend = 7.6, yend = 2.3), arrow = arrow(length = unit(0.01, "npc")), curvature = 0) +
    annotate("text", 7.99, 3.5, hjust = 0, 
           label = "75th Percentile") +
  geom_curve(aes(x = 8, y = 3, xend = 7.905, yend = "Hazel"), arrow = arrow(length = unit(0.01, "npc")), curvature = -.2) +
  
  guides(size = "none") +
  coord_cartesian(xlim = c(7.4, 8.4)) +
  labs(x = "Episode Ratings",
       y = "Characters",
       title = "The five main characters do not seem to have different distributions of\nepisode ratings",
       caption = "\nData from IMDb.com  |  @PhilliPhresh")
```

I'm not surprised Dr. Spaceman is stably near the top of the list.
