---
title: "Web Scraping 30 Rock Data"
author: "Phillip Sanderell"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages

```{r}
library(tidyverse)
library(rvest)
library(here)
```

# Scrape raw data from IMDb

There are seven seasons of 30 Rock.

```{r}
rock_raw <- tibble()

for (s in 1:7) {
  
  season_link <- paste0("https://www.imdb.com/title/tt0496424/episodes?season=", s)
  
  season_page <- read_html(season_link)
  
  se_index <- 
    season_page %>% 
    html_elements(".zero-z-index div") %>% 
    html_text(trim = TRUE)
  
  ep_name <- 
    season_page %>% 
    html_elements("#episodes_content strong a") %>% 
    html_text(trim = TRUE)
  
  ep_link <- 
    season_page %>% 
    html_elements("#episodes_content strong a") %>% 
    html_attr("href") %>% 
    paste0("https://www.imdb.com", .)
  
  air_date <- 
    season_page %>% 
    html_elements(".airdate") %>% 
    html_text(trim = TRUE)
  
  ep_desc <- 
    season_page %>% 
    html_elements(".item_description") %>% 
    html_text(trim = TRUE)
  
  no_votes <- 
    season_page %>% 
    html_elements(".ipl-rating-star__total-votes") %>% 
    html_text(trim = TRUE)
  
  rating <- 
    season_page %>% 
    html_elements(".ipl-rating-star.small .ipl-rating-star__rating") %>% 
    html_text(trim = TRUE)
  
  rock_raw <- 
    rbind(rock_raw, 
          tibble(se_index, ep_name, ep_link, air_date, ep_desc, no_votes, rating))
  
  #It's always polite to sleep between accessing pages, even if not asked on the robots.txt
  Sys.sleep(3 + runif(1, 1, 2))
}
```

## Pull directors and writers for each episode---Try 1

```{r}
ep_link <- "https://www.imdb.com/title/tt1635816/?ref_=ttep_ep4"

ep_page <- read_html(ep_link)

writers <-
  ep_page %>% 
  html_elements("div.PrincipalCredits__PrincipalCreditsPanelWideScreen-hdn81t-0.iGxbgr > ul > li:nth-child(2) a") %>%
  html_text %>% 
  paste(collapse = "__")

directors <- 
  ep_page %>% 
  html_elements("div.PrincipalCredits__PrincipalCreditsPanelWideScreen-hdn81t-0.iGxbgr > ul > li:nth-child(1) a") %>%
  html_text %>% 
  paste(collapse = "__")
```

The method above currently has some flaws:

-   If an episode has more writers or directors than fit on the page, it does not display all of them.

-   Tina Fey created the show, so she's listed on every episode as a writer even if she didn't really write the script. Sure I could collect her in every episode and then remove her but then I might remove her when she actually is the writer of the script.

I've got a different idea I want to try for collecting writers and directors. It seems like each episode page has a cleaner, more detailed list of credits where writers as listed as (created by), (written by), (story editor), (executive story editor), etc. If I can get the link for each episode's, which there might be a pattern for that I can apply to the links I already have, then I could try looping through those pages to pull only the writers and directors I'm actually interested in.

## Pull directors and writers for each episode---Try 2

The chunk below gets links for each episode's credits page. This seems to work when I spot check some of the links.

```{r}
# The block below gets links for each episode's credits page
credits <- 
  rock_raw %>% 
  mutate(ep_credit_link = str_extract(ep_link, ".+?\\?")) %>% 
  mutate(ep_credit_link = str_replace(ep_credit_link, "\\?", "fullcredits?ref_=tt_ov_wr_sm")) %>% 
  select(se_index, ep_credit_link)
```

![]()

```{r}
get_credits <- function(link) {

  credit_page <- read_html(link)
  
  directors <-
    credit_page %>% 
    html_element("#director+ .simpleCreditsTable") %>% 
    html_table() %>% 
    # filter(grepl("(directed by)", X3)) %>% 
    pull(X1) %>% 
    paste(collapse = "__")
  
  writers <- 
    credit_page %>% 
    html_element("#writer+ .simpleCreditsTable") %>% 
    html_table() %>% 
    filter(grepl("(written by)", X3)) %>% 
    pull(X1) %>% 
    paste(collapse = "__")
  
  Sys.sleep(1 + runif(1, 1, 2))
  
  data.frame(directors = directors, writers = writers) %>% 
    return()
}

# This cycles through all episodes, could take a while, ~6 minutes.
dw <- sapply(credits$ep_credit_link, 
             FUN = get_credits, 
             USE.NAMES = FALSE)

# Put dw output into a neat tibble.
directors <- 
  dw[1, ] %>% 
  tibble(directors = .) %>% 
  unnest(directors)

writers <- 
  dw[2, ] %>% 
  tibble(writers = .) %>% 
  unnest(writers)

dw_credits <- 
  credits %>%
  cbind(directors, writers)
```

Nice! Looks like this method worked well.

# Combining and cleaning the data

Our original data set and credits data set need combined. Some variables need cleaned.

```{r}
rock <-
  full_join(rock_raw, dw_credits, by = "se_index") %>% 
  separate(se_index, into = c("season", "ep_no")) %>% 
  mutate(season = parse_number(season),
         ep_no = parse_number(ep_no)) %>% 
  mutate(air_date = str_remove(air_date, "\\."),
         air_date = parse_date(air_date, format = "%d %b %Y"),
         no_votes = parse_number(no_votes),
         rating = parse_number(rating))
```

Check the data for missing values.

```{r}
rock %>% View
```

From a quick glance, looks pretty good.

# Export data

I considered exporting the data as a .Rds because I would like to make sure R saved the variable classes for when I import the data in a clean session, but then it wouldn't be accessible if someone wanted to import it into, for example, Excel or Python, later. So I decided to just export as .csv and hope the clean format of the dates and numbers help R default to the write classes, if they don't, oh well.

```{r}
write_csv(rock, here("rock.csv"))
```
